---
title: "RMED901-Exam Report-Group 5"
author: Alexander Vietheer, Dinastry Pramadita Zakiudin, Marta Espevold Hjelmeland
  and Shanshan Xu
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
[Check the document](https://mitt.uib.no/groups/68014/files/folder/DATA?preview=4450357)

This is the **group 5 exam report** for the course **RMED901 22H / Data science with R for medical researchers**, we used the data set from the paper in the New England Journal of Medicine, entitled, [A Randomized Trial of Rectal Indomethacin to Prevent Post-ERCP Pancreatitis, pages 1414-1422 volume 366, 2012 edition, authored by the Elmunzer, BJ, Higgins PDR, et al.](https://www.nejm.org/doi/full/10.1056/NEJMoa1111103). 



#Loading the package for analysis
```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(here)
```
**1.Read the data set**
```{r}
myData <- read_delim(here("DATA", "exam_nontidy.txt"), delim = "\t")
```
**2. Clean the data set**

**#Data cleaning 1:** We observe some variables starting with number,and the variable "feature type" has space between the variable name, so we rename these by using the pipe-rename.
```{r}
myData <- myData %>% 
  rename(Dose_asa_81 = `81asa`,
         Dose_asa_325 = `325asa`,
         feature_type = `feature type`)
```
**#Data cleaning 2:** We observed the variable "id" has the two parts, we checked the codebook of the dataset, the first integer 1-4 indicates site of the study so we use separate() function to separate the id column
```{r}
myData <- myData %>% 
  separate(col = id, 
           into = c("site", "id"), 
           sep = "_")
```
**#Data cleaning 3:** Remove the duplicated row in the dataset using the distinct() function (there are 10 rows with duplicates)
```{r}
myData <- myData %>% 
  distinct()
```
**#Data cleaning 4:** Then we pivot the "feature_type" column into sod and pep columns
```{r}
myData$feature_type <- as.factor(myData$feature_type)
myData$feature_value <- as.numeric(myData$feature_value)
myData <- myData %>% 
  pivot_wider(names_from = `feature_type`, values_from = feature_value)
```
**#Data cleaning 5:** Check the missing data
```{r}
naniar::gg_miss_var(myData) 
```

**#Data cleaning 6:** Remove unnecessary columns from your dataframe: acinar, train, amp, pdstent. Use subset() fuction to delete column by name
```{r}
df=subset(myData, select = -c(acinar, train, amp, pdstent))
```

#df is the data frame excluded the cloumns:acinar, train, amp, pdstent from the day 6 task.

**3. Change the variable types:** Make necessary changes in variable types

#A column showing whether age is higher than 35 or not: values High/Low
```{r}
myData <- myData %>% 
  mutate(agegroup = case_when(age <= 35 ~ "Low",
                              age > 35 ~ "High"))
```
#A numeric column showing risk as a percentage of highest possible risk (5.5)
```{r}
myData$risk <- as.numeric(myData$risk)
myData <- myData %>% 
  mutate(riskpercentage = risk / 5.5)
```
#A column showing pep as No/Yes
```{r}
myData <- 
  myData %>% 
  mutate(Newpep = pep, Newpep = if_else(Newpep == "0", "No", "Yes"))
```
#A numeric column showing multiplication of age and risk for each person
```{r}
myData$age<-as.numeric(myData$age)
myData <- myData %>% 
  mutate(AgemultiplybyRisk = age*risk )
```
**4. Change the order of the column as id, site, age and other columns; and arrange id column of the dataset in the order of increasing number**
```{r}
myData <- myData %>% 
  select(id, everything())
arrange(myData, id)
```
**5. Combine the two data sets.**

#Read the additional dataset
```{r}
antibodyData <- read_delim(here("DATA", "exam_joindata.txt"), delim = "\t")
```
#Separate the columns "id" in the "joindata" dataset.
```{r}
antibodyData <- antibodyData %>% 
  separate(col = id, 
           into = c("site", "id"), 
           sep = "_")
```
#Combine the two datasets using pipe.
```{r}
Fulldataset <- myData %>% 
  left_join(antibodyData, by = c("id","site"))
```
#The Fulldataset contains n = 602 observations and 37 variables. We will work on this Fulldataset.

#Exploring the data. We observed there are many missing data (402) from antibodyData especially in female
```{r}
is.na(Fulldataset)
naniar::gg_miss_var((Fulldataset), facet = gender)
```



#Group the data by gender. We had 476 females and 126 males, this could be why there was a big difference in missing data between the genders.
```{r}
Fulldataset %>% 
  group_by(gender) %>% 
  count()
```

#Stratify the data by a categorical column and a report of min, max, mean and sd of a numeric column.
```{r}
Fulldataset %>% 
  group_by(gender) %>% 
  summarise(min(risk, na.rm = T),max(risk, na.rm = T), mean(risk, na.rm = T), sd(risk, na.rm = T))
```

## Data cleaning 3: remove the duplicated row in the dataset using the distinct() function}
```{r}
myData<- myData %>% 
  distinct()
```
### there are 10 rows with duplicates final n = 1204 

## Tidy 4: Then we pivot the column into sod and pep columns
```{r}
myData$feature_type <- as.factor(myData$feature_type)
myData$feature_value <- as.numeric(myData$feature_value)
myData <- myData %>% 
  pivot_wider(names_from = `feature_type`, values_from = feature_value)
```

#Stratify the data using pipe by a categorical column and a report of min, max, mean and sd of a numeric column for a defined set of observations
```{r}
Fulldataset %>% 
  group_by(gender) %>% 
  summarise(min(antibody, na.rm = T),max(antibody, na.rm = T), mean(antibody, na.rm = T), sd(antibody, na.rm = T))
```


#Stratify the data only for persons with no recurrent pancreatitis
```{r}
Fulldataset %>% 
  filter(recpanc == 0)
```

#Stratify the data only for persons recruited in University of Kentucky
```{r}
Fulldataset %>% 
  filter(site == 3)
```

#Startify the data only for persons older than 45
```{r}
Fulldataset %>% 
  filter(age > 45)
```

#Stratify the data only for persons with risk higher than 2 and sphincter of Oddi dysfunction type 2. We observed no patients fall in this category
```{r}
Fulldataset %>% 
  filter(risk > 2 & sod_type ==2)
```

#Explore the data by creating a table to observe gender and sphincter of Oddi dysfunction type 
```{r}
Fulldataset %>% 
  select(gender, sod_type) %>% 
  count(gender, sod_type)
```

## adding extra libraries
```{r}
library("patchwork")
library("ggplot2")
library("corrplot")
```



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
