---
title: "RMED901-Exam Report-Group 5"
author: Alexander Vietheer, Dinastry Pramadita Zakiudin, Marta Espevold Hjelmeland
  and Shanshan Xu
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
[Check the document](https://mitt.uib.no/groups/68014/files/folder/DATA?preview=4450357)

This is an **exam report** for the course **RMED901 22H / Data science with R for medical researchers**, we used the datset from the paper in the New England Journal of Medicine, entitled, A Randomized Trial of Rectal Indomethacin to Prevent Post-ERCP Pancreatitis, pages 1414-1422 volume 366, 2012 edition, authored by the Elmunzer, BJ, Higgins PDR, et al. More details of the data set can be found in the link above.

```{r}
#read the data
myData <- read_delim(here("DATA", "exam_nontidy.txt"), delim = "\t")
```


```{r}
#Loading the package for analysis
library(tidyverse)
library(ggplot2)
library(dplyr)
library(here)
```

```{r}
#Data cleaning 1: We observe some variables starting with numbers,and the variable "feature type" has space between the variable name, so we rename these by using the pipe-rename.
myData <- myData %>% 
  rename(Dose_asa_81 = `81asa`,
         Dose_asa_325 = `325asa`,
         feature_type = `feature type`)
```

```{r}
#Data cleaning 2: We observed the variable "id" has the two parts, we checked the codebook of the dataset, that the first integer 1-4 indicates site of the study so we use separate() function to separate the id column
myData <- myData %>% 
  separate(col = id, 
           into = c("site", "id"), 
           sep = "_")
```

```{r}
#Data cleaning 3: remove the duplicated row in the dataset using the distinct() function
myData<- myData %>% 
  distinct()
# there are 10 rows with duplicates final n = 1204 
```

```{r}

```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
